### 2021-10-22 README.Rmd
gifs

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
```

## Overview

ggplot2 is a system for declaratively creating graphics, based on [The Grammar of Graphics][gg-book]. You provide the data, tell ggplot2 how to map variables to aesthetics, what graphical primitives to use, and it takes care of the details.

## Installation

### Glenelg region

```{r glenelg, echo = FALSE}
knitr::include_graphics("figs/glenelg_predator_activity.gif")
```

### Otway ranges

```{r otways, echo = FALSE}
knitr::include_graphics("figs/otways_predator_activity.gif")
```