### 2021-12-26 README.Rmd
rm ind

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
```

## Overview

This repository contains the data and code of my fourth PhD data chapter.

Here I've built generalised additive models ('GAMs'), using the ['mgcv' R-package](https://cran.r-project.org/web/packages/mgcv/mgcv.pdf), of invasive predator spatiotemporal activity in south-west Victoria, Australia.

This paper aims to provide inference on the drivers of invasive predator spatiotemporal activity in temperate, heterogeneous landscapes. Particularly whether a subordinate mesopredator (feral cat *Felis catus*) exhibits (signs of) behavioural avoidance of an apex predator (Red fox *Vulpes vulpes*).

There are three model types, differing in how the hourly smooth was specified:

1. Predator activity across space (tensor product of hour and coordinates)
2. Predator activity across vegetation types (hierarchical model of hour x vegetation type)
3. Feral cat activity potential avoidance of foxes (tensor product of hour and coordinates)

## Acknowledgements

The idea to use GAMs for this chapter came from Prof. Nick Golding - Nick also wrote me some initial simulation code and helped me get started with GAMs. I also learnt most of these methods and gained inspiration from [online course materials](https://noamross.github.io/mgcv-esa-2018/) and [this paper on heirarchical GAMs](https://peerj.com/articles/6876/) from Noam Ross, Gavin Simpson, David Miller and Eric Pedersen.

I compiled camera-trap data from three distinct studies:

- My PhD surveys 2017 - 2019.

Funding and support for these surveys came from the University of Melbourne, Conservation Ecology Centre, the Victorian  Government Department of Environment, Land, Water, and Planning, Parks Victoria, Arthur Rylah Institute for Environmental Research, Parks Victoria.

- Glenelg Ark monitoring data 2013 - 2019.

Robley, A., Moloney, P., Stringer, L. and Donald, S. (2020). Glenelg Ark 2005–2019: long-term predator and native mammal response to predator control. Arthur Rylah Institute for Environmental Research Technical Report Series No. 318. Department of Environment, Land, Water and Planning, Heidelberg, Victoria.

- Otway Ark monitoring data 2016 - 2018.

Robley, A., Fanson, B. and Parks Victoria West Coast District Team. (2019). The Otway Ark: response of predators and native species 2016–2018. Arthur Rylah Institute for Environmental Research Technical Report Series No. 299. Department of Environment, Land, Water and Planning, Heidelberg, Victoria.

```{r photo, echo = FALSE}
knitr::include_graphics("figs/steph.jpg")
```

**Surveys were carried out the traditional lands of the Gunditjmara and Eastern Maar.**

## Repository structure

**Raw data is contained in the "raw_data/" folder**.

- *"predator_records.csv"* has a row for every cat and fox camera-trap detection (no time separation between consecutive detections). 
- *"camdata.csv"* contains camera-trap deployment information and explanatory variables. Note that there are plenty of camera-traps which were deployed but did not detect predators - predator_records and camdata need to be merged to account for absences.

**R scripts are contained in the "r_scripts/" folder**.

- script and function to fit lorelograms using adapted code from [Iannarilli et al. 2019](https://doi.org/10.1111/2041-210X.13308) are contained in the *"/lorelograms"* folder. This was an exploratory analysis to choose the best time-to-indepence for consecutive detections of species on each camera-trap (I chose 30 mins). This involves creating a minute by minute detection history so is computationally intensive.
- *"format_records_hour_gams.R"* reformats the raw data files into a usable dataframe for spatiotemporal GAMs. This accounts for solar times using [average double anchorage methods](https://doi.org/10.1111/2041-210X.13290), removes non-indepedent detections (within 30 mins) and reshapes into a dataframe where, for each camera-trap, there are 0 - 23 rows for hour of the day, a column for cat and fox filled with the number of 'independent' detections in each hour, as well as columns for deployment information and explanatory variables. This dataframe is saved at **"derived_data/predator_counts_hour.csv"**. 
- *"models.R"* uses mgcv to fit GAMs
- scripts beginning with *"plot_"* contain code used to create the figures. These scripts require the *"models"* script to be run first in the same session. 
- *"individual_cats_reformat_model_plots.R"* reformats "cat_id_detections.csv" into a dataframe, runs heirarchical GAMs for each cat / cat category, and creates the figures (not included in the paper).

**Figures for the manuscript are contained in the "figs/" folder**.

## GIFs of Invasive predator spatiotemporal activity:

#### Glenelg region

```{r glenelg, echo = FALSE}
knitr::include_graphics("figs/glenelg_predator_activity.gif")
```

#### Otway Ranges

```{r otways, echo = FALSE}
knitr::include_graphics("figs/otways_predator_activity.gif")
```